Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-08 manuscript/appA.Rmd
updates appendix with IPM model dscriptions

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

@@

Appendix A: Detailed model descriptions
=======================================
### Tredennick and Adler

Here we describe the statistical models in complete detail, including full description of the Bayesian models and Stan code for each model. We write each model as three parts: (1) a data model (likelihood), (2) a process model, and (3) a parameter model. We then combine these three models to write the posterior and joint distributions as a single model statement. Following the full model expression we provide the Stan code that directly corresponds the statement of posterior and joint distributions. In our expressions of the posterior and joint distributions we use general probability notation where [.] represents some unspecified distribution. We do this for clarity of model presentation. For each model we include a table of prior distributions for necessary parameters. Also for clarity, we avoid subscripting species (the *j* subscripts in the main text).

\newpage{}

Individual level models
-----------------------
### Survival
Data, process, and parameter models.

\begin{align}
&\mathtt{data} \qquad &\textbf{y} \sim \text{Bernoulli}(\boldsymbol{\mu}) \\
&\mathtt{process} \qquad &\text{logit}(\mu_{iQ,t}) = \gamma_{t} + \phi_{Q} + \beta_{t}x_{ij,t} + \omega w_{i,t} + \theta_{k}C_{k,t} \\
&\mathtt{parameters} \qquad &\gamma_{t} \sim \text{Normal}(\tilde{\gamma}, \sigma_\gamma^2)\\
& \qquad & \phi_{Q} \sim \text{Normal}(0,\sigma_Q^2) \\
& \qquad & \beta_{t} \sim \text{Normal}(\tilde{\beta}, \sigma_\beta^2) \\
& \qquad & \omega \sim \text{Normal}(0, 0.0001) \\
& \qquad & \boldsymbol{\theta} \sim \text{Normal}(0, 0.0001)
\end{align}

Full expression of posterior and joint distributions.
\begin{align}
[\boldsymbol{\gamma}, \tilde{\gamma}, \boldsymbol{\phi}, \boldsymbol{\beta}, \tilde{\beta}, \omega, \boldsymbol{\theta}, \sigma_Q^2, \sigma_\beta^2] &\propto \\
&\prod_{t=1}^T \prod_{i=1}^n [y_{iQ,t} | \gamma_{t}, \phi_{Q}, \beta_{t}, \omega, \boldsymbol{\theta}] [\gamma_{t} | \tilde{\gamma}] [\beta_{t} | \tilde{\beta}, \sigma_\beta^2] \times \\
&\prod_{Q=1}^{Q_{tot}} [\phi_{Q} | \sigma_Q^2] \times \\
&[\boldsymbol{\theta}] [\omega] [\tilde{\gamma}] [\tilde{\beta}] [\sigma_Q^2] [\sigma_\beta^2]
\end{align}

Stan code for model.
```{r survival_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0,upper=1> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- inv_logit(a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + 
                         crowdEff[n] + climEff[n]);
  }
}
model{
  // Priors
  a_mu ~ uniform(-300,300);
  w ~ uniform(-100,100);
  b1_mu ~ uniform(-100,100);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ uniform(-10,10);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);

@@

>>>>

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

@@

Appendix A: Detailed model descriptions
=======================================
Here we describe the statistical models in complete detail, including full description of the Bayesian models and Stan code for each model. We write each model as three parts: (1) a data model (likelihood), (2) a process model, and (3) a parameter model. We then combine these three models to write the posterior and joint distributions as a single model statement. Following the full model expression we provide the Stan code that directly corresponds the statement of posterior and joint distributions. In our expressions of the posterior and joint distributions we use general probability notation where [.] represents some unspecified distribution. We do this for clarity of model presentation. For each model we include a table of prior distributions for necessary parameters. Also for clarity, we avoid subscripting species (the *j* subscripts) after initial description of each model.

Survival (IPM)
--------------
We used logistic regression to model survival probability ($S$) of genet $i$ from species $j$ in quadrat group $Q$ from time $t$ to $t+1$:

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat group location, and $\theta^{S}_{k}$ is the fixed parameter for the effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). Note that the vector of climate covariates (**C**) includes climate variable interactions and climate$\times$size interactions. We include density-dependence by estimating the effect of crowding on the focal individual by other individuals of the same species. $\omega$ is the effect of crowding and $w_{t,Q}$ is the crowding experienced by the focal individual at time $t$ in quadrat group $Q$. We include a size$\times$crowding interaction effect ($\nu^{S}$).

**Data, process, and parameter models**

\begin{align}
&\mathtt{data} \qquad &\textbf{y} \sim \text{Bernoulli}(\boldsymbol{\mu}) \\
&\mathtt{process} \qquad &\text{logit}(\mu_{iQ,t}) = \gamma_{t} + \phi_{Q} + \beta_{t}x_{i,t} + \omega w_{i,t} + \nu w_{i,t}x_{i,t} + \theta_{k}C_{k,t} \\
&\mathtt{parameters} \qquad &\gamma_{t} \sim \text{Normal}(\tilde{\gamma}, \sigma_\gamma^2)\\
& \qquad & \phi_{Q} \sim \text{Normal}(0,\sigma_Q^2) \\
& \qquad & \beta_{t} \sim \text{Normal}(\tilde{\beta}, \sigma_\beta^2) \\
& \qquad & \omega \sim \text{Uniform}(-100, 100) \\
& \qquad & \nu \sim \text{Uniform}(-100, 100) \\
& \qquad & \boldsymbol{\theta} \sim \text{Uniform}(-10, 10) \\
& \qquad & \tilde{\gamma} \sim \text{Uniform}(-300, 300) \\
& \qquad & \tilde{\beta} \sim \text{Uniform}(-100, 100) \\
& \qquad & \sigma_Q^2 \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_\gamma^2 \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_\beta^2 \sim \text{Cauchy}(0, 5)
\end{align}

**Full expression of posterior and joint distributions**
\begin{align}
[\boldsymbol{\gamma}, \tilde{\gamma}, \boldsymbol{\phi}, \boldsymbol{\beta}, \tilde{\beta}, \omega, \boldsymbol{\theta}, \sigma_Q^2, \sigma_\beta^2] &\propto \\
&\prod_{t=1}^T \prod_{i=1}^n [y_{iQ,t} | \gamma_{t}, \phi_{Q}, \beta_{t}, \omega, \boldsymbol{\theta}] [\gamma_{t} | \tilde{\gamma}] [\beta_{t} | \tilde{\beta}, \sigma_\beta^2] \times \\
&\prod_{Q=1}^{Q_{tot}} [\phi_{Q} | \sigma_Q^2] \times \\
&[\boldsymbol{\theta}] [\omega] [\tilde{\gamma}] [\tilde{\beta}] [\sigma_Q^2] [\sigma_\beta^2]
\end{align}

**Stan code for model**
```{r survival_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0,upper=1> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- inv_logit(a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + 
                         crowdEff[n] + climEff[n]);
  }
}
model{
  // Priors
  a_mu ~ uniform(-300,300);
  w ~ uniform(-100,100);
  b1_mu ~ uniform(-100,100);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ uniform(-10,10);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);

@@

Growth (IPM)
------------
We modeled growth as Gaussian process describing genet size at time $t+1$ as a function of size at $t$ and climate covariates:

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
Y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \sigma_{j})
\end{align}

where $x$ is log genet size and all other parameters are as described for the survival regression.

**Data, process, and parameter models.**
\begin{align}
&\mathtt{data} \qquad &\textbf{y} \sim \text{Normal}(\boldsymbol{\mu}, \boldsymbol{\sigma}^2) \\
&\mathtt{process} \qquad &\mu_{iQ,t+1} = \gamma_{t} + \phi_{Q} + \beta_{t}x_{i,t} + \omega w_{i,t} + \nu w_{i,t}x_{i,t} + \theta_{k}C_{k,t} \\
& \qquad & \sigma^2_{iQ,t} = ae^{b\mu_{iQ,t+1}} \\
&\mathtt{parameters} \qquad &\gamma_{t} \sim \text{Normal}(\tilde{\gamma}, \sigma_\gamma^2)\\
& \qquad & \phi_{Q} \sim \text{Normal}(0,\sigma_Q^2) \\
& \qquad & \beta_{t} \sim \text{Normal}(\tilde{\beta}, \sigma_\beta^2) \\
& \qquad & \omega \sim \text{Normal}(0, 0.001) \\
& \qquad & \nu \sim \text{Normal}(0, 0.001) \\
& \qquad & \boldsymbol{\theta} \sim \text{Normal}(0, 0.001) \\
& \qquad & \tilde{\gamma} \sim \text{Normal}(0, 0.001) \\
& \qquad & \tilde{\beta} \sim \text{Normal}(0, 0.001) \\
& \qquad & a \sim \text{Normal}(0, 1000) \\
& \qquad & b \sim \text{Normal}(0, 1000) \\
& \qquad & \sigma_Q^2 \sim \text{Uniform}(0, 1000) \\
& \qquad & \sigma_\gamma^2 \sim \text{Uniform}(0, 1000) \\
& \qquad & \sigma_\beta^2 \sim \text{Uniform}(0, 1000)
\end{align}

**Full expression of posterior and joint distributions**
\begin{align}
[\boldsymbol{\gamma}, \tilde{\gamma}, \boldsymbol{\phi}, \boldsymbol{\beta}, \tilde{\beta}, \omega, \boldsymbol{\theta}, \sigma^2, \sigma_Q^2, \sigma_\beta^2] &\propto \\
&\prod_{t=1}^T \prod_{i=1}^n [y_{iQ,t} | \gamma_{t}, \phi_{Q}, \beta_{t}, \omega, \boldsymbol{\theta}, \sigma^2] [\gamma_{t} | \tilde{\gamma}] [\beta_{t} | \tilde{\beta}, \sigma_\beta^2] \times \\
&\prod_{Q=1}^{Q_{tot}} [\phi_{Q} | \sigma_Q^2] \times \\
&[\boldsymbol{\theta}] [\omega] [\tilde{\gamma}] [\tilde{\beta}] [\sigma_Q^2] [\sigma_\beta^2]
\end{align}

**Stan code for model**
```{r grow_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  vector[N] Y; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  real gint[G];
  real tau;
  real tauSize;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  real mu[N];
  real<lower=0> sigma[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + crowdEff[n] + climEff[n];
    sigma[n] <- sqrt((fmax(tau*exp(tauSize*mu[n]), 0.0000001)));  
  }
}
model{
  // Priors
  a_mu ~ normal(0,1000);
  w ~ normal(0,1000);
  b1_mu ~ normal(0,1000);
  tau ~ normal(0,1000);
  tauSize ~ normal(0,1000);
  sig_a ~ uniform(0,1000);
  sig_b1 ~ uniform(0,1000);
  sig_G ~ uniform(0,1000);
  for(g in 1:G)
      gint[g] ~ normal(0, sig_G);
  for(c in 1:Covs)
    b2[c] ~ normal(0,1000);
  for(y in 1:Yrs){
    a[y] ~ normal(a_mu, sig_a);
    b1[y] ~ normal(b1_mu, sig_b1);
  }

  // Likelihood
  Y ~ normal(mu, sigma);
}
```

Recruitment (IPM)
-----------------
Our data allows us to track new recruits, but we cannot assign a specific parent to new genets. So, for recruitment, we work at the quadrat level and model the number of new individuals of species $j$ in quadrat $q$ recruiting at time $t+1$ as a function of quadrat "effective cover" ($A'$) in the previous year ($t$). Effective cover is a mixture of observed cover ($A$) in the focal quadrat ($q$) and the mean cover across the entire group ($\bar{A}$) of $Q$ quadrats in which $q$ is located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where $p$ is a mixing fraction between 0 and 1 that is estimated within the model.

We assume the number of individuals, $Y^{R}$, recruiting at time $t+1$ follows a negative binomial distribution:

\begin{equation}
Y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where $\lambda$ is the mean intensity and $\zeta$ is the size parameter. We define $\lambda$ as:

\begin{equation}
\lambda_{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where $A'$ is effective cover ($\text{cm}^2$) of species $j$ in quadrat $q$ and all other terms are as in the survival and growth regressions.

**Data, process, and parameter models**
\begin{align}
&\mathtt{data} \qquad &\textbf{y} \sim \text{NegBin}(\boldsymbol{\lambda},\zeta) \\
&\mathtt{process} \qquad &\lambda_{jq,t+1} = A'_{q,t}e^{(\gamma_{t} + \phi_{Q} + \theta_{k}C_{k,t} + \omega\sqrt{A'_{q,t}})} \\
&\mathtt{parameters} \qquad &\gamma_{t} \sim \text{Normal}(\tilde{\gamma}, \sigma_\gamma^2)\\
& \qquad & \phi_{Q} \sim \text{Normal}(0,\sigma_Q^2) \\
& \qquad & \omega \sim \text{Uniform}(-100, 100) \\
& \qquad & \boldsymbol{\theta} \sim \text{Uniform}(-10, 10) \\
& \qquad & \tilde{\gamma} \sim \text{Uniform}(-300, 300) \\
& \qquad & \sigma_Q^2 \sim \text{Cauchy}(0, 5) \\
& \qquad & \sigma_\gamma^2 \sim \text{Cauchy}(0, 5)
\end{align}

**Full expression of posterior and joint distributions**
\begin{align}
[\boldsymbol{\gamma}, \tilde{\gamma}, \boldsymbol{\phi}, \boldsymbol{\beta}, \tilde{\beta}, \omega, \boldsymbol{\theta}, \sigma_Q^2, \sigma_\beta^2] &\propto \\
&\prod_{t=1}^T \prod_{i=1}^n [y_{iQ,t} | \gamma_{t}, \phi_{Q}, \beta_{t}, \omega, \boldsymbol{\theta}] [\gamma_{t} | \tilde{\gamma}] [\beta_{t} | \tilde{\beta}, \sigma_\beta^2] \times \\
&\prod_{Q=1}^{Q_{tot}} [\phi_{Q} | \sigma_Q^2] \times \\
&[\boldsymbol{\theta}] [\omega] [\tilde{\gamma}] [\tilde{\beta}] [\sigma_Q^2] [\sigma_\beta^2]
\end{align}

**Stan code for model**
```{r rec_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] parents1; // crowding vector
  vector[N] parents2; // crowding vector
}
parameters{
  real a_mu;
  vector[Yrs] a;
  vector[Covs] b2;
  real dd;
  real gint[G];
  real<lower=0> sig_a;
  real<lower=0> theta;
  real<lower=0> sig_G;
  real<lower=0, upper=1> u;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  vector[N] trueP1;
  vector[N] trueP2;
  vector[N] lambda;
  vector[N] q;
  climEff <- C*b2;
  for(n in 1:N){
    trueP1[n] <- parents1[n]*u + parents2[n]*(1-u);
    trueP2[n] <- sqrt(trueP1[n]);
    mu[n] <- exp(a[yid[n]] + gint[gid[n]] + dd*trueP2[n] + climEff[n]);
    lambda[n] <- trueP1[n]*mu[n];
    q[n] <- lambda[n]*theta;
  }
}
model{
  // Priors
  u ~ uniform(0,1);
  theta ~ uniform(0,100);
  a_mu ~ normal(0,1000);
  dd ~ uniform(-100,100);
  sig_a ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  for(g in 1:G)
      gint[g] ~ normal(0, sig_G);
  for(c in 1:Covs)
    b2 ~ uniform(-100,100);
  for(y in 1:Yrs){
    a[y] ~ normal(a_mu, sig_a);
  }

  // Likelihood
  Y ~ neg_binomial_2(q, theta);
}
```

### 2015-07-08 manuscript/manuscript.Rmd
updates appendix with IPM model dscriptions

@@

Third, ecologists as a community need to get serious about reporting uncertainty. There is a strong culture around explicitly considering model uncertainty, but parameter uncertainty is often ignored. In some cases this is because the easiest statistical methods do no make propogating parameter uncertainty easy. But even using Bayesian approaches that allow straightforward integration of model fitting and forecasting [@Hooten2014; @Hobbs2015] is not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts cen we begin to produce better ones.

>>>>

These results lead us to conclude that detailed demographic data is necessary to forecast climate change impacts on plant populations. This is unwelcome news since this data is difficult to collect and the models built on such data are of little use to land managers that make decisions at scales beyond that of traditional population models [@Queenborough2011]. While density-structured approaches may fail when climate covariates are considered, there are other alternatives. For example, @Clark2011a use Forest Inventory and Analysis (FIA) data to parameterize a population model with multiple vital rates and climate dependence. Another example are distributed efforts like PlantPopNet (http://plantago.plantpopnet.com) that will allow researchers to estimate variation around climate responses for widespread species. Lastly, we foresee new approaches on the horizon that leverage photo/video of plots and advanced object recognition algorithms to streamline plant mapping and digitizing efforts.

@@

Third, ecologists as a community need to get serious about reporting uncertainty. There is a strong culture around explicitly considering model uncertainty, but parameter uncertainty is often ignored. In some cases this is because the easiest statistical methods do no make propogating parameter uncertainty easy. Even using Bayesian approaches that allow straightforward integration of model fitting and forecasting [@Hobbs2015] is not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts can we begin to produce better ones.

### 2015-07-08 manuscript/manuscript.tex
updates appendix with IPM model dscriptions

\newpage{}

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit a single regression model for population growth. We describe the
statistical models separately since fitting the models required
different approaches. All models contain five climate covariate that we
chose \emph{a priori}: ``water year'' precipitation at \emph{t}-1
(lagppt); fall through spring precipitation at \emph{t}-1 and \emph{t}-2
(ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1
and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}
is the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions.

@@

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a
year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope
parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat
group location, and $\theta^{S}_{k}$ is the fixed parameter for the
effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). We
include density-dependence by estimating the effect of crowding on the
focal individual by other individuals of the same species. $\omega$ is
the effect of crowding and $w_{t,Q}$ is the crowding experienced by the
focal individual at time $t$ in quadrat group $Q$.

@@

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. 3). The log normal likelihood
includes a truncation (T{[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

@@

Third, ecologists as a community need to get serious about reporting
uncertainty. There is a strong culture around explicitly considering
model uncertainty, but parameter uncertainty is often ignored. In some
cases this is because the easiest statistical methods do no make
propogating parameter uncertainty easy. But even using Bayesian
approaches that allow straightforward integration of model fitting and
forecasting (Hooten and Hobbs 2014, Hobbs and Hooten 2015) is not simple
when using modeling approaches like integral projection models that
separate the model fitting and simulation stages (Rees and Ellner 2009).
However, as we have done here, it is still possible to include parameter
uncertainty by drawing parameter values from MCMC iterations, taking
care to draw all parameters from the same chain and iteration to account
for their correlations. Only by being honest about our forecasts cen we
begin to produce better ones.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison of population models based on individual and
aggregated data in a climate change context. Our results confirm
theoretical arguments (Clark et al. 2011b) and empirical evidence (Clark
et al. 2011a, 2012) that individual responses are critical to predicting
species' responses to climate change. Thus, forecasts from aggregate
level models should be viewed with caution and should never be
unaccompianed by uncertainty.

@@

Hooten, M. B., and N. T. Hobbs. 2014. A guide to Bayesian model
selection for ecologists. Ecological Monographs 85:3--28.

>>>>

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit a single regression model for population growth. We describe the
statistical models separately since fitting the models required
different approaches. All models contain five climate covariate that we
chose \emph{a priori}: ``water year'' precipitation at \emph{t}-1
(lagppt); fall through spring precipitation at \emph{t}-1 and \emph{t}-2
(ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1
and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}
is the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions. Climate $\times$ size interactions are for climate
main effects only, that is we do not include interactions among size and
interacting climate effects.

@@

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

where $x_{ij,t}$ is the log of genet size, $\gamma^{S}_{j,t}$ is a
year-specific intercept, $\beta^{S}_{j,t}$ is the year-specific slope
parameter for size, $\phi^{S}_{jQ}$ is the random effect of quadrat
group location, and $\theta^{S}_{k}$ is the fixed parameter for the
effect of the $k$th climate covariate at time $t$ ($C_{k,t}$). Note that
the vector of climate covariates (\textbf{C}) includes climate variable
interactions and climate$\times$size interactions. We include
density-dependence by estimating the effect of crowding on the focal
individual by other individuals of the same species. $\omega$ is the
effect of crowding and $w_{t,Q}$ is the crowding experienced by the
focal individual at time $t$ in quadrat group $Q$. We include a
size$\times$crowding interaction effect ($\nu^{S}$).

@@

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. 3). Again, note that the climate
covariate vector (\textbf{C}) includes the climate$\times$cover
interaction. The log normal likelihood includes a truncation
(T{[}0,1{]}) to ensure that predicted values do not exceed 100\% cover.

@@

These results lead us to conclude that detailed demographic data is
necessary to forecast climate change impacts on plant populations. This
is unwelcome news since this data is difficult to collect and the models
built on such data are of little use to land managers that make
decisions at scales beyond that of traditional population models
(Queenborough et al. 2011). While density-structured approaches may fail
when climate covariates are considered, there are other alternatives.
For example, ({\textbf{???}}) use Forest Inventory and Analysis (FIA)
data to parameterize a population model with multiple vital rates and
climate dependence. Another example are distributed efforts like
PlantPopNet (\url{http://plantago.plantpopnet.com}) that will allow
researchers to estimate variation around climate responses for
widespread species. Lastly, we foresee new approaches on the horizon
that leverage photo/video of plots and advanced object recognition
algorithms to streamline plant mapping and digitizing efforts.

@@

Third, ecologists as a community need to get serious about reporting
uncertainty. There is a strong culture around explicitly considering
model uncertainty, but parameter uncertainty is often ignored. In some
cases this is because the easiest statistical methods do no make
propogating parameter uncertainty easy. Even using Bayesian approaches
that allow straightforward integration of model fitting and forecasting
(Hobbs and Hooten 2015) is not simple when using modeling approaches
like integral projection models that separate the model fitting and
simulation stages (Rees and Ellner 2009). However, as we have done here,
it is still possible to include parameter uncertainty by drawing
parameter values from MCMC iterations, taking care to draw all
parameters from the same chain and iteration to account for their
correlations. Only by being honest about our forecasts can we begin to
produce better ones.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated data in a climate change context. Our results
confirm theoretical arguments (Clark et al. 2011b) and empirical
evidence (Clark et al. 2011a, 2012) that individual responses are
critical to predicting species' responses to climate change. Thus,
forecasts from aggregate level models should be viewed with caution and
should never be unaccompianed by uncertainty. Given the importance of
demographic data and its current difficulty to collect, we advocate for
research on modern methods to collect demographic data more efficiently
across environmental gradients in space and time. For example, recent
advances in object recognition algorithms (Liu et al. 2014) provide new
opportunities to speed up data collection and digital processing.

@@

Liu, Y., Y. Jang, W. Woo, and T.-K. Kim. 2014. Video-Based Object
Recognition Using Novel Set-of-Sets Representations.
